PhD Title \\ PhD Thesis

### 2024-06-07 Latex_files/Thesis/chapters/chapter 2/main.tex
more updates...

Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser. From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle. The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations. The differential forms of which are given below:
\begin{align}
  \nabla \cdot \mathbf{E}
  &= \frac{\rho_ v}{\epsilon_0}
  \\
  \nabla \cdot \mathbf{B}
  &= 0
  \\
  \nabla \times \mathbf{E}
  &= \frac{\partial \mathbf{B}}{\partial t}
  \\
  \nabla \times \mathbf{B}
  &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}    
\end{align}
where $\rho_v$ is the free charge density, ${\bf j}_0$ is the charge
current, $\epsilon_0$ is the permittivity of free space.

@@

\begin{align}
	E_{inc}(r) &= E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r) \\
	E_{int}(r) &= E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right) \\
	E_{scat}(r) &= E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
\end{align}

Where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user. However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on. From an optical trapping perspective the force exerted by a focused electric field can be found by computing the maxwell stress tensor which only requires the total magnitude of the incident and scattered fields, essentially computing the momentum contained in the incident beam and how much of it has been transferred to the target particle.

Lornez-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory. The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions. In the latter case we can simplify the scattering problem by approximating our target as a single dipole, focusing only its interactions with the incident field.

@@

\section{Simulation of spherical aggregates}

@@

	\frac{\mathbf{d}\vec{r}(t)}{\mathbf{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + \sqrt{2\vec{D}_x}\eta(t)

@@

	\frac{\mathbf{d}\vec{u}(t)}{\mathbf{dt}} = \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t) + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)

@@

		\frac{\bold{d}\vec{r}(t)}{\bold{dt}} &= \frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2D}\eta(t) \\
		\frac{\bold{d}\vec{u}(t)}{\bold{dt}} &= \frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t) + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)

@@

Fortunately, we do not need to solve these directly as the latter two random variables can be easily approximated if the thermal energy of the system is know, and the rate of change can be assumed as linear if we take a sufficiently small time step that $\Delta t~\ll~\kappa_x/\gamma \ \& \ \Delta t \ll \kappa_u/\gamma_R$. In doing so we now only need to compute the optical force and torque applied to our dimer, this has already been compiled in a MATLAB package called \textit{Optical Tweezer Toolbox} or \textit{ott} \cite{Nieminen2007}. In which they use the results from \cite{Farsund1996} to compute both the optical force and torque using the beam coefficients - the form given by \cite{Crichton2000THEMD} is an easier form to compute: 

@@

	\bold{F_z}=&-\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{n,m}b*_{n,m}-p_{n,m}q*_{n,m}] \right. \\ 
	&+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}  \\

@@

	\bold{T_z}=&-\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{n,m}|^2+|b_{n,m}|^2-|p_{n,m}|^2-|q_{n,m}|^2] \right. \\ 
	&+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}  \\

@@

Where $a_{n,m}$, $b_{n,m}$, $p_{n,m}$ $q_{n,m}$ are the beam coefficients of the incident and scattered fields respectively. We can get the transverse force and torque components in a similar form by applying a simple rotation transformation. In order to get the scattered beam coefficients we can use \textit{MSTM} \cite{Mackowski2011} to calculate the T-matrix of our dimer (or any spherical aggregate). Vigilante et al compiled together a python package that combines both \textit{ott} and \textit{MSTM} to simulate the behaviour of spherical aggregates within a predefined optical trap.

@@

\subsection{Simulated Quadrant Photodiode}
In order to simulate a typical experimental set up with a QPD installed as a position detection system we need to evaluate the total magnitude of the electric field incident on the photo-diode surface. While trapping a micro-particle, the scattered and incident fields combine together and interfere with one another. These fields are collected by a condenser lens in the far field limit and are focused onto the QPD surface, the total intensity can be evaluated as:
\begin{align}
I(x,y) = \epsilon_0c\left|
\begin{bmatrix} 
	E_{i,x}(x,y)+E_{s,x}(x,y) \\ 
	E_{i,y}(x,y)+E_{s,y}(x,y) \\ 
	E_{i,z}(x,y)+E_{s,z}(x,y)
\end{bmatrix} \right|^2 \times step(NA_c-\sqrt{x^2+y^2})
\end{align}
The last term is simply a representative step term that defines the outer limit by which we evaluate the electric field, this is analogous to our condenser lens removing noise from other light sources by only accepting light at a specific acceptance angle defined by its numerical aperture $NA_c$. Depending on the relative size of our particle we can adjust the acceptance angle, this has very little effect on the transverse signals, but for axial evaluations of a trapped particle the numerical aperture should be tuned so that the resultant response curve has negative slope in order to allow for axial position detection, the method for finding this angle $\theta_\Theta$ is discussed in \cite{Friedrich2012}.

The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). \textit{Ott} uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:

\begin{align}
	E_{inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
\end{align}

Where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, \textit{ott} allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD. For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully. In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.

>>>>

Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser. From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle.  The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations. The differential forms of which are given below
\cite{Jackson_1975}:
\begin{align}
  \nabla \cdot \mathbf{E}
  &= \frac{\rho_ v}{\epsilon_0}
  \\
  \nabla \cdot \mathbf{B}
  &= 0
  \\
  \nabla \times \mathbf{E}
  &= \frac{\partial \mathbf{B}}{\partial t}
  \\
  \nabla \times \mathbf{B}
  &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}    
\end{align}
where $\rho_v$ is the free charge density, ${\bf j}_0$ is the charge
current, and $\epsilon_0$ is the permittivity of free space.

@@

\begin{align}
  E_{\rm inc}(r)
  &=
    E_0 \sum_{n=1}^\infty a_{n}M_{1n}^{(1)}(r)+b_{n}N_{1n}^{(1)}(r)
  \\
  E_{\rm int}(r)
  &=
    E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left(-id_{n}N_{1n}^{(1)}(r)+c_{n}M_{1n}^{(1)}(r)\right)
  \\
  E_{scat}(r)
  &=
    E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left(-ip_{n}N_{1n}^{(3)}(r)+q_{n}M_{1n}^{(3)}(r)\right)
\end{align}
where $a_n$, $b_n$, $c_n$, $d_n$, $p_n$, and $q_n$ are the expansion coefficients of each of the fields, for the incident field computing its expansion coefficients is possible via analytical methods and are completely dependent on the beam conditions imposed by the user. However, computing the expansion coefficients of the internal and external fields be far more tedious depending on the shape of the target and what properties of the scattered field are desired - all of which is discussed later on. From an optical trapping perspective the force exerted by a focused electric field can be found by computing the maxwell stress tensor which only requires the total magnitude of the incident and scattered fields, essentially computing the momentum contained in the incident beam and how much of it has been transferred to the target particle.

Lorenz-Mie theory can be applied to describe the scattering from any particle regardless of size, however as we either increase or decrease the size of the target particles the infinite series converges allowing one to ignore much of the tedium of Lorenz-Mie theory. The ray optics model is what is achieved when the size of the target particle far exceeds that of the laser wavelength and thus individual ray's can have independent contributions. In the latter case we can simplify the scattering problem by approximating our target as a single dipole, focusing only its interactions with the incident field.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation of spherical aggregates}

@@

\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + \sqrt{2\vec{D}_x}\eta(t)

@@

  \frac{{d}\vec{u}(t)}{{dt}}
  =
  \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t)
  + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)

@@

  \frac{{d}\vec{r}(t)}{{dt}}
  &=
    \frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2D}\eta(t)
  \\
  \frac{{d}\vec{u}(t)}{{dt}}
  &=
    \frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t)
    + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)

@@

Fortunately, we do not need to solve these directly as the latter two
random variables can be easily approximated if the thermal energy of
the system is know, and the rate of change can be assumed as linear if
we take a sufficiently small time step that
$\Delta t~\ll~\kappa_x/\gamma \ \& \ \Delta t \ll
\kappa_u/\gamma_R$. In doing so we now only need to compute the
optical force and torque applied to our dimer, this has already been
compiled in a MATLAB package called \textit{Optical Tweezer Toolbox}
or \textit{ott} \cite{Nieminen2007}. In which they use the results
from \cite{Farsund1996} to compute both the optical force and torque
using the beam coefficients - the form given by
\cite{Crichton2000THEMD} is an easier form to compute:

@@

  \bold{F_z}
  &=
    -\frac{1}{4\pi k^2}\sum_{n,m} \left(\frac{m}{n(n+1)}\Im[a_{n,m}b*_{n,m}-p_{n,m}q*_{n,m}] \right.
  \\ 
  &+\frac{1}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
  \\

@@

  \bold{T_z}
  &=
    -\frac{1}{8\pi k^3}\sum_{n,m} \left(\frac{m}{n(n+1)}[|a_{n,m}|^2+|b_{n,m}|^2-|p_{n,m}|^2-|q_{n,m}|^2] \right.
  \\ 
  &+\frac{2}{n+1}\left[\frac{n(n+2)(n-m+1)(n+m+1)}{(2n+1)(2n+3)} \right]^{1/2}
  \\

@@

where $a_{nm}$, $b_{nm}$, $p_{nm}$, and $q_{nm}$ are the beam coefficients of the incident and scattered fields respectively. We can get the transverse force and torque components in a similar form by applying a simple rotation transformation. In order to get the scattered beam coefficients we can use \textit{MSTM} \cite{Mackowski2011} to calculate the T-matrix of our dimer (or any spherical aggregate). Vigilante et al compiled together a python package that combines both \textit{ott} and \textit{MSTM} to simulate the behaviour of spherical aggregates within a predefined optical trap.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Simulated Quadrant Photodiode}

In order to simulate a typical experimental set up with a QPD installed as a position detection system we need to evaluate the total magnitude of the electric field incident on the photo-diode surface. While trapping a micro-particle, the scattered and incident fields combine together and interfere with one another. These fields are collected by a condenser lens in the far field limit and are focused onto the QPD surface, the total intensity can be evaluated as:
\begin{align}
I(x,y) = \epsilon_0c\left|
\begin{bmatrix} 
	E_{i,x}(x,y)+E_{s,x}(x,y) \\ 
	E_{i,y}(x,y)+E_{s,y}(x,y) \\ 
	E_{i,z}(x,y)+E_{s,z}(x,y)
\end{bmatrix} \right|^2 \times step(NA_c-\sqrt{x^2+y^2})
\end{align}
The last term is simply a representative step term that defines the outer limit by which we evaluate the electric field, this is analogous to our condenser lens removing noise from other light sources by only accepting light at a specific acceptance angle defined by its numerical aperture $NA_c$. Depending on the relative size of our particle we can adjust the acceptance angle, this has very little effect on the transverse signals, but for axial evaluations of a trapped particle the numerical aperture should be tuned so that the resultant response curve has negative slope in order to allow for axial position detection, the method for finding this angle $\theta_\Theta$ is discussed in \cite{Friedrich2012}.

The incident beam is simple enough to define given our set up parameters, for the sake of simplicity we assume that our beam is a Laguerre-Gaussian beam of mode $[0.0, 0.0]$ (which is simply a pure Gaussian beam). \textit{Ott} uses a point matching approach to approximate the beam shape coefficients of the incident field by fitting it to the far field estimate, the beam is of the form:
\begin{align}
	E_{\rm inc}(kr)=\sum^\infty_n\sum^n_{m=-n}a_{mn}RgM_{nm}(kr)+b_{nm}RgN_{nm}(kr)
\end{align}
where $RgM_{nm}(kr)$ \& $RgN_{nm}(kr)$ are regular vector spherical wave functions, \textit{ott} allows us to change the basis of the the incident beam to suit our needs, because we are measuring in the far field we want to set our incident beam to be an outgoing spherical wave so that we can compute the intensity on the QPD. For spherical waves the field can either be expressed as an incoming/outgoing wave (with a singularity at the origin) or as a regular wave around the origin; for incoming/outgoing waves the wave functions use the first/second forms of the Hankel function respectfully. In order to compute the regular spherical wave at the origin we replace the Hankel function with the Bessel function which is simply the average of the first and second forms of the Hankel function, so at the origin we avoid a singularity of the EM field.
